### 2022-06-19 latex/Chapters/Chapter2.tex

% Chapter Template

\chapter{Games} % Main chapter title

\label{Chapter2} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Sliding Puzzle}

blablabla

%-----------------------------------
%	SUBSECTION 1
%-----------------------------------
\section{Rubiks' Cube}

blabla