### 2022-06-19 latex/Chapters/Chapter2.tex
update

% Chapter Template

\chapter{Games} % Main chapter title

\label{sec:Games} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Sliding Puzzle}

The first puzzle I will focus on is the sliding puzzle (see \cite{SlidingPuzzle}).

%-----------------------------------
%	SUBSECTION 1.1
%-----------------------------------
\subsection{Search Space \& Solvability}

blabla

%-----------------------------------
%	SUBSECTION 1.2
%-----------------------------------
\subsection{Optimal Cost}

blabla

%-----------------------------------
%	SECTION 2
%-----------------------------------
\section{Rubiks' Cube}

blabla