### 2022-07-31 latex/Chapters/Code.tex
update

% Chapter Template

\chapter{Code} % Main chapter title

\label{Chapter3} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

%\section{Code organisation}

The code I have developed for this project is all publicly available on my github page (\cite{FB}). It can easily be installed using the setup file provided, which makes it easy to then use Python's customary import command to play with the code.
The code is organised in several sub modules and makes use of factories in plenty of places so that I can easily try out different puzzles, dimensions, search techniques, heuristics, network architecture, etc... without having to change anything except the parameters passed in the command line. Here is a visual overview of the code base with the main dependencies between the main submodules and classes. Solid arrows indicate inheritance (e.g. AStar inherits from SearchStrategy), while dotted lines indicate usage (e.g. AStar uses Heuristic, DeepReinforcementLearner uses DeepLearning, etc..).

\begin{landscape}
\begin{figure}[H]
\centering
\includegraphics[scale=0.49]{./Figures/codebase}
%\decoRule
\caption[Codebase]{Code base}
\label{fig:Codebase}
\end{figure}
\end{landscape}

Let me now describe what each submodule does in more details:

\section{rubiks.core}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebasecore}
%\decoRule
\caption[Codebase]{rubiks.core}
\label{fig:Codebasecore}
\end{figure}
This submodule contains base classes that make the code base easier to use, debug, and extend. It contains the following:
\begin{itemize}
\item \textbf{Loggable}: a wrapper around Python's logger which automatically picks up classes' names at init and format things (dict, series and dataframes in particular) in a nicer way.
\item \textbf{Parsable}: a wrapper around ArgumentParser, which allows to construct objects in the project from command line, to define dependencies between object's configurations and to help a bit with typing of configs. The end result is that you can pretty much pass **kw\_args everywhere and it just works.
\item \textbf{Factory}: a typical factory pattern. Concrete factories can just define what widget they produce and the factory will help construct them from **kw\_args (or command line, since Factory inherits from Parsable)
\end{itemize}

\section{rubiks.puzzle}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebasepuzzle}
%\decoRule
\caption[Codebase]{rubiks.puzzle}
\label{fig:Codebasepuzzle}
\end{figure}
This submodule contains:
\begin{itemize}
\item \textbf{Puzzle}: a Factory of puzzles. It defines states and actions in the abstract, and provides useful functions to apply moves, shuffle, generate training sets, tell if a state is the goal, etc. Puzzle can manufacture the two following types of puzzles:
\item \textbf{SlidingPuzzle}. Implements the states and moves of the sliding puzzle.
\item \textbf{RubiksCube}. Implements the states and moves of the Rubik's cube.
\\
\\
In addition, this module contains a \textbf{Puzzled} base class which most classes below inherit from. That allow e.g. heuristics, search algorithms, solvers and learners to know what puzzle and dimension they operate on, without having to reimplement these basic facts in each of them.
\end{itemize}

\section{rubiks.search}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebasesearch}
%\decoRule
\caption[Codebase]{rubiks.search}
\label{fig:Codebasesearch}
\end{figure}
This modules contains graph search strategies. I have actually reused the code I implemented for one of the AIPnT assignments here. It contains the following classes:
\begin{itemize}
\item \textbf{Node}: which contains the state of a graph, as well as link to the previous (parent) state, action that leads from the latter to the former and the cost of the path so far.
\item \textbf{SearchStrategy}, a Factory class which can instantiate the following three types of search strategies to find a path to a goal:
\item \textbf{BreadthFirstSearch}, which is obviously an optimal strategy, but not particularly efficient.
\item \textbf{DepthFirstSearch}, which is not an optimal strategy, and also generally not particularly efficient.
\item \textbf{AStar}, which is optimal, and as efficient as the heuristic it makes use of is.
\end{itemize}

\section{rubiks.heuristics}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebaseheuristics}
%\decoRule
\caption[Codebase]{rubiks.heuristics}
\label{fig:Codebaseheuristics}
\end{figure}
\label{HSS}
This module contains base class Heuristic, also a Factory. Heuristic can instantiate the following heuristics, which we can use in the AStar strategy from the previous section:
\begin{itemize}
\item \textbf{Manhattan}: This heuristic is specific to the SP. It simply adds for each tile (other than the empty tile) the $L_{1}$ distance between their current position and their target position. We can quickly see that this heuristic is admissible. Indeeed, each move in the SP moves one (and one only) tile by one position. In other puzzles, it is generally the case that a move will affect the position of many tiles (e.g. RC). Therefore, if tiles could somehow freely move on top of one-another (that is, we remove the constraint that there can be at most one tile per compartment, the number of moves necessary to solve the SP would exactly be the Manhattan distance.
\\
\\
The reason why it is interesting to have a known admissible heuristic is that we can obvioulsy compare other heuristics (DL, DRL, DQL, etc) in terms of optimality.
\\
\\
I have also implemented an improvement to the Manhattan distance, which I shall call Manhattan++ in here (and in code logs and graphs) and can be activated by simply passing \textit{plus=True} to the Manhattan Heuristic (see example \ref{ASSS} as well as a thorougher performance comparison on the (n=2, m=5) SP in \ref{MHComp}). It is based on the concept of linear constraints that I read about in lecture notes from Washington University (\cite{SlidingPuzzleLectureNotes}). The idea is simply that when tiles are already placed on their target row or column, but not in the expected order, we will need to get the tiles out of the way of one another to get to the goal state, and this is not accounted for by the Manhattan distance. For instance, in the following (n=2, m=3) SP, the Manhattan distance does not account for the fact that, at best, tile \red 3 \black needs to get out of the way for tiles \blue 1 \black and \blue 2 \black to move, and then needs to get back in its row, adding a cost of 2 to the Manhattan distance.

\begin{center}
\begin{five}
\setrow{2}{\red 3, \blue 1, \blue 2}
\setrow{1}{4, 5, 0}
\end{five}
\end{center}

Two important things to notice are that linear constraints across rows and columns (which I might generically refer to as \textit{lines} in the following) can to be added without breaking admissibility (hence giving more pessimistic, or accurate, cost estimates than simple Manhattan). This is because if a tile is involved in two linear constraints, it will need to get out of the way both horizontally and vertically. The second thing is that when several pairs in a line are not in order, we cannot simply add 2 for each distinct out-of-order pair, the right penalty to add is more subtle than that and needs to be computed recursively. For instance, let us now consider the following configuration:
\begin{center}
\begin{five}
\setrow{2}{\red 3, \red 2, \red 1}
\setrow{1}{4, 5, 0}
\end{five}
\end{center}
The correct penalty to add is not 6 (3 times 2 since all pairs (1, 2), (1, 3) and (2, 3) are out of order, but only 4. Indeed if, say, tile 3 got somehow out of the way at the back of the SP (imagine just another dimension there where we can move tiles) and tile 2 got out of the way by moving down to let tile 1 pass across, we could be done by simply adding 4 to the Manhattan distance (2 to move tile 3 out and back, and 2 to move tile 2 out and back). The correct way to compute the penalty cost for linear constraints is therefore to do it recursively, taking the minimum additional cost of moving either left-most or right-most tile of the line under consideration out of the way (that additional cost to move these left-most or right-most tile is 2 if not at their expected order in the line, 0 otherwise) plus the penalty of reordering the rest of the line.
\\
\\
Finally, as suggested by the reference lecture notes (which give very vague details about the above subtleties), I have precomputed all the penalties for all possible rows, columns and all possible tiles ordering they could have and saved the corresponding penalties in a database. For memory efficiency, I also only saved penalties which are non-zero. The very first time any call to Manhattan++ is made, for a given dimension (n, m), the appropriate data base is computed and populated.
\\
\\
Notice that for an (n, m) SP, there are n rows, each of which can have $\frac{(n * m)!}{(n * m - m)!}$ different ordering of tiles and m columns which can have n columns, each can have $\frac{(n * m)!}{(n * m - n)!}$ different ordering of tiles. This means the pre-computations and data-base sizes for the Manhattan++ heuristic are actually manageable, as it grows much slower than the number of possible puzzles. Taking into account that we only store non-zero penalties, for instance the (n=4, m=4) Manhattan++ database only has  records, and the (n=5, m=5) has .

\item \textbf{PerfectHeuristic}: this reads from a data base the optimal costs, pre-computed by the PerfectLearner (see below \ref{PLcode})
\item \textbf{DeepLearningHeuristic}: this uses a network which has been trained using DRL by the DeepReinforcementLearner (see below \ref{DRLcode})
\end{itemize}

\section{rubiks.deeplearning}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebasedeeplearning}
%\decoRule
\caption[Codebase]{rubiks.deeplearning}
\label{fig:Codebasedeeplearning}
\end{figure}
This module is a wrapper around Pytorch. It contains:
\begin{itemize}
\item \textbf{DeepLearning}: a Puzzled Loggable Factory that can instantiate some configurable deep networks, and provide the necessary glue with the rest of the code base so that puzzles be seemlessly passed to the networks and trained on.
\item \textbf{FullyConnected}: wrapper around a Pytorch fully connected network, with configurable layers and size.
\item \textbf{Convolutional}: tdb
\end{itemize}

\section{rubiks.learners}
\label{sec:codelearners}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebaselearners}
%\decoRule
\caption[Codebase]{rubiks.learners}
\label{fig:Codebaselearners}
\end{figure}
\label{PLcode}
\label{DRLcode}
This module implements learners, which learn something from a puzzle, store what they learnt, and can display interesting things about what they learnt.

\begin{itemize}
\item \textbf{Learner} is a Puzzled Loggable Factory. It provides some common code to learners (to save or purge what they learnt), kick off learning and plot results. Concrete derived implementation define what and how they learn, and what interesting they can display about this learning process. Currently the two implemented learners are:
\item \textbf{PerfectLearner}: It instantiates an optimal solver ($A^{*}$ with a configurable heuristic - but will only accept heuristic that advertise themselves as optimal. The learning consists in generating all the possible configuration of the considered puzzle, solve them with the optimal solver, and save the optimal cost of it as well as those of the whole solution path. The code allows for parallelization, stop and restart so that we can run on several different occasions and keep completing a database of solutions if necessary or desired. Once the PerfectLearner has completed its job, it can display some interesting information, such as the puzzle's God's number, the distribution of number of puzzles versus optimal cost, the hardest configuration it came across, and how long it took it to come up with the full knowledge of that puzzle. I will show in section \ref{PLSS} how to run an example. Notice that for puzzles of too high dimension, where my computing resources will not allow to solve exhaustively all the configurations of a given dimension, this class can still be used to populate a data base of optimal costs, which can then be used by DeepLearner. If it is to be used this way, the PerfectLearner can be configured to use perfectly random configurations to learn from, rather than going through the configurations one by one in a well defined order.

\item \textbf{DeepLearner} tbd
\item \textbf{DeepReinforcementLearner}: It instantiates a DeepLearning (network), and trains it using DRL. It then saves the trained network, which can then be used in the DeepLearningHeuristic we have seen earlier in section \ref{HSS}. The

\end{itemize}

\section{rubiks.solvers}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebasesolvers}
%\decoRule
\caption[Codebase]{rubiks.solvers}
\label{fig:Codebasesolvers}
\end{figure}
This module implements solvers, which solve puzzles. The base class Solver is a Factory of solvers, and in addition to being able to instantiating the following types of solvers, can run different solvers through a similal sequences of random puzzles (for various increasing degrees of difficulty (shuffling), and/or perfectly shuffled ones) and display a comparison of how they perform in a number of metrics.

\begin{itemize}
\item \textbf{DFSSolver} TBD
\item \textbf{BFSSolver} TBD
\item \textbf{AStarSolver} TBD
\item \textbf{NaiveSlidingSolver} TBD
\item \textbf{MonteCarloSearchTreeSolver} TBI I want to implement this in August when working on the Rubiks' ... following the Rubik's paper fromMcAleer \& Agostilenni et al \cite{https://doi.org/10.48550/arxiv.1805.07470}
\end{itemize}

\section{rubiks.scripts}

\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebasescripts}
%\decoRule
\caption[Codebase]{rubiks.scripts}
\label{fig:Codebasescripts}
\end{figure}

Finally it is worth noting that the code will save on disk a lot of data (e.g. the learners will save what they have learnt, e.g. a Pytorch network or a data base of optimal costs, the performance comparison will run solvers versus very many configurations of puzzles and save the results for later being able to display) etc... The base of the tree to save all this data can be chosen by setting up the "RUBIKSDATA" environment variable. If not, it will go somewhere in you HOME :)