### 2022-09-03 latex/Appendices/Kociemba.tex
update

% Chapter Template

\chapter{Kociemba Bug} % Main chapter title

\label{sec:Kociemba} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

Discuss annoyance with Kociemba 3x3x3 with respect to color scheme

%-----------------------------------
%	SECTION 1
%-----------------------------------