### 2022-09-07 latex/RHULMScAIFrancoisBerrier.tex

\documentclass[11pt]{article} % do not change this line
\input{BigDataStyle.txt}      % do not change this line
\input{Packages/Packages}

\addbibresource{references.bib} % The filename of the bibliography
\newcommand{\chapter}{\section}
\newcommand{\Section}{\subsection}
\newcommand{\Subsection}{\subsubsection}

\hypersetup{%
  colorlinks = true,
  linkcolor  = black
}

\emergencystretch=5mm
\tolerance=400
\allowdisplaybreaks[4]

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{problem}[theorem]{Problem}

\theoremstyle{definition}
\newtheorem*{remark}{Remark}

\title{Solving \\ the \\ 15-Puzzle \\ \& \\ Rubik's Cube}
\author{Fran\c{c}ois Berrier}

\newcommand{\Programme}{Artificial Intelligence}

\hypersetup{%
  citecolor=mdtRed
}

\begin{document}
\maketitle

\begin{abstract}
\input{Abstract/Abstract}
\end{abstract}

\noindent
\textbf{Acknowledgements}
\newline
\newline
\input{Acknowledgements/Acknowledgements}

%\tableofcontents % Prints the main table of contents

\newpage

\hypersetup{%
  colorlinks = true,
  linkcolor  = orange
}
\input{Chapters/Objectives}
\include{Chapters/DRLSearch}

\printbibheading
\hypersetup{pdfpagemode={UseOutlines},
bookmarksopen=true,
bookmarksopenlevel=0,
hypertexnames=false,
colorlinks=true,% Set to false to disable coloring links
citecolor=magenta,% The color of citations
linkcolor=red,% The color of references to document elements (sections, figures, etc)
urlcolor=mdtRed,% The color of hyperlinks (URLs)
pdfstartview={FitV},
unicode,
breaklinks=true,
}
\printbibliography[type=article,heading=subbibliography,title={Journal Papers}]
\printbibliography[type=book,heading=subbibliography,title={Books}]
\printbibliography[type=misc,heading=subbibliography,title={Misc}]
\end{document}