### 2022-09-08 latex/Chapters/Conclusion.tex
update

% Chapter Template

\chapter{Conclusion \& Professional Issues} % Main chapter title

\label{Conclusion} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\Section{Learning}
Learnt about puzzles, and also even took the time to learn a couple of sequences of moves that are sufficient to solve \textbf{RC}, 2x2x2 in labout one minute, 3x3x3 in about 2 minutes.

\Section{Implementation}
Implemented a bunch of solvers. Ideally should do in C++ for speed and multithreaded (esp for MCTS) but time was limited so I prioritized breadth over depth.

\Section{Deep Reinforcement/Q Learning}
Quite impressive, without too much effort, how well they perform. However, I feel to really get more out of them (higher dimensions in particular) would need careful tuning of the convergence criteria.

\Section{Professional issues}
See \cite{ManyWorlds}
See \cite{https://doi.org/10.1111/1758-5899.12718} and \cite{Bostrom2014}