### 2022-09-12 latex/Abstract/Abstract.tex
trimming abstract and ack

\label{Abstract}
\noindent \textbf{Motivation} -- Reinforcement Learning (\textbf{RL}), exposed with brilliant clarity in \cite{Sutton1998}, has until recently known less success than we might have hoped for. Its framework is very appealing and intuitive. In particular, the mathematical beauty of Value and Q iteration (\cite{WatkinsThesis}) for discrete state and action spaces, blindly iterating from \textit{any} initial value, is quite profound. Sadly, it had until recently proven hard to achieve practical success with these methods. Inspired however by the seminal work of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have made a lot of progress towards designing algorithms capable of learning and solving, \textit{without human knowledge}, the Rubik's Cube (\textbf{RC}) by using Deep Q-Learning (\textbf{DQL})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}). I will attempt to implement a variety of solvers combining $A^{*}$ and heuristics, some handcrafted, others trained on sequences of puzzles using Deep Learning (\textbf{DL}), \textbf{DRL} or \textbf{DQL}, to solve the 15-puzzle and the Rubik's cube.
\\
\\
\textbf{Organisation of this thesis} -- In section 1, I discuss what I hope to get out of this project, both in terms of personal learning, as well as in terms of tangible results (solving some puzzles!). In section 2, I do a recap of the methods used in the project. Section 3 is dedicated to the mathematics of the sliding puzzle \textbf{SP} and the \textbf{RC}. I give then an overview of my implementation in section 4. Sections 6 and 7 present my results on respectively the \textbf{SP} and the \textbf{RC} and section 8 concludes. A couple of appendices complement the main text: the first one details how to use the code base, the second describes a limitation I bumped into while using the kociemba library.
\\
\\
\noindent \textbf{Acknowledgements} -- 
\input{Acknowledgements/Acknowledgements}