### 2022-09-12 latex/Acknowledgements/Acknowledgements.tex
spell check

I am grateful to Bank of America for sponsoring my part time 2-year MSc in Artificial Intelligence at Royal Holloway and would like to thank my managers Mitrajit Dutta and Stephen Thompson for their support. The company of my colleague Saurabh Kumar, as he embarked on the same MSc journey and with whom I have had countless debates about the new cool stuff we learnt, has made the whole process much more enjoyable.
\\
I would also like to thank Professor Chris Watkins for supervising my project. It was an honour to attend his lectures on \textbf{DL} during the 2020-21 school year, and I tried to absorb some of his wisdom at our meetings during this project. I am glad to have followed his advice at our last meeting: I stopped shaving for a month, which did free enough time for me to implement \textbf{DQL} (joint policy and value learning), as well as an \textbf{MCTS} algorithm, to nicely complement my assortment of solvers.
\\
Finally, my wife Thanh Nh\~a is my biggest supporter in everything I do, and her continued encouragement in my pursuit of this MSc has made it so much easier to juggle between a demanding banking job, running 50 to 70 miles a week in preparation for the Chicago marathon, and the coursework, lectures and exam revisions.